Distinguishing NFW and Isothermal Density Profiles with Weak Gravitational Lensing

### 2018-05-07 paper.tex
new references and comparison figure, more paper writing

\graphicspath{ {figures/} }

@@

\begin{abstract}
We examine the feasibility of distinguishing NFW and cored isothermal density profiles using weak gravitational lensing shear.
 To do so, we use Python to model lenses in the two different profiles, as well as background galaxies to be lensed.
 Analyzing the simulated data of these lensed galaxies gives us insight into how we can distinguish the differences between the two profiles.
 This method is expected to be helpful in the analysis of real observation data in the future.
\end{abstract}

\section{Introduction}
Gravitational lensing allows us to investigate the matter density distributions of cosmic objects by observing the characteristic distortions imparted on other objects in the background. The principle is applicable on a large range of scales, from all-sky mass maps based on cosmic microwave background shear to studies of individual galaxies and clusters.
A common use of lensing is

Two commonly used density profiles in weak gravitational lensing are the isothermal profile and the Navarro-Frenk-White (NFW) profile.
We start with introducing the general characteristics of spherical density profiles, followed by cored isothermal and NFW profiles.
We then describe the purpose of our project, to distinguish between NFW and isothermal profiles, in detail.
The goal of this project is to devise a method to analyze lensed galaxy cluster data and find which density profile is more probable between the isothermal and NFW profiles.
In order to do so, we generate simulated data sets and analyze them to find the more probable profiles they would fit into.
This analysis method is expected to be usable in determining the characteristics of real observed data as well.

@@

@@

The critical surface density marks the typical boundary between what is considered strong and weak lensing:
\begin{equation}
\Sigma_\mathrm{crit} = \frac{c^2}{4\pi G} \frac{D_S}{D_{SL} D_L}
\end{equation}

@@

\section{Cored Isothermal Sphere Profile}
The cored isothermal sphere (CIS) profile is related to the singular isothermal sphere, which is often used to describe halos and other collapsed astrophysical objects because of its simple formulation. Unlike the singular isothermal sphere, the cored isothermal sphere does not have a density singularity at its center due to a finite core radius $r_c$. The CIS density profile is defined:

@@

\begin{figure}
    \includegraphics{isothermalproperties.pdf}
    \caption{Lensing quantities for a CIS profile with $M_{200} = 10^{15} M_\odot$ and $r_c = 10$ kpc.}
    \label{}
\end{figure}

@@

\begin{figure}
    \includegraphics{nfwproperties.pdf}
    \caption{Lensing quantities for a NFW profile with $M_{200} = 10^{15} M_\odot$ and $c = 10$.}
    \label{}
\end{figure}

@@

\begin{enumerate}
    \item Consider single foreground lens halo with many background galaxies.
    \begin{itemize}
        \item Start with one NFW halo, then maybe consider more tests.
    \end{itemize}
    \item Construct background galaxies:
    \begin{itemize}
        \item Number density on sky: 50 galaxies/square arcminute
        \item Intrinsic ellipticity drawn from Gaussian distribution with $\mu=0, \sigma=0.2$
        \item Assume they are all that the same distance $D_S$ since this can be determined by redshift (neglecting some noise)
    \end{itemize}
    \item Recommended values:
    \begin{itemize}
        \item $z_L = 0.3$ (bullet cluster)
        \item $z_S = 1.0$ (Hubble Deep Field)
        \item $M_{halo} = 10^{15} M_\odot$
    \end{itemize}
\end{enumerate}

@@

\begin{figure}
    \includegraphics[width=0.49\textwidth]{isothermalellipticities.pdf}
    \includegraphics[width=0.49\textwidth]{nfwellipticities.pdf}
    \caption{plots}
    \label{}
\end{figure}

@@

\section{Results and Conclusion}

@@

>>>>

\graphicspath{ {figures/} }
\linespread{1.3}

@@

\begin{abstract}
We examine the feasibility of distinguishing NFW and cored isothermal density profiles using weak gravitational lensing shear.

We model lenses in the two different profiles, as well as background galaxies to be lensed.
Analyzing the simulated data of these lensed galaxies gives us insight into how we can distinguish the differences between the two profiles.
This method is expected to be helpful in the analysis of real observation data in the future.
\end{abstract}

\section{Introduction}
Gravitational lensing allows us to investigate the matter density distributions of cosmic objects by observing the characteristic distortions imparted on other objects in the background. The principle is applicable on a large range of scales, from all-sky mass maps based on cosmic microwave background shear to studies of individual galaxies and clusters. In particular, by observing the coherent distortions of many background galaxies, the shear profile, and consequently the density profile of a foreground halo lens may be determined.

Two commonly used density profiles in the field are the isothermal profile and the Navarro-Frenk-White (NFW) profile. The NFW profile has been found to fit simulated dark matter halos very well \citep{}, while the isothermal profile .
Both profiles have infinite extent

Lensing offers one of the only ways to probe the distribution
In this paper, we will investigate the distinguishability of
given galaxy-galaxy lensing data

We start with introducing the general characteristics of spherical density profiles, followed by cored isothermal and NFW profiles.
We then describe the purpose of our project, to distinguish between NFW and isothermal profiles, in detail.
The goal of this project is to devise a method to analyze lensed galaxy cluster data and find which density profile is more probable between the isothermal and NFW profiles.
In order to do so, we generate simulated data sets and analyze them to find the more probable profiles they would fit into.
This analysis method is expected to be usable in determining the characteristics of real observed data as well.

@@

The lens halo is located at an angular diameter distance $D_L$ from the observer and we want to consider its effects on a source object behind it at an angular diameter distance $D_S$.

@@

The critical surface density marks the typical boundary between what is considered strong and weak lensing:
\begin{equation}
\Sigma_\mathrm{crit} = \frac{c^2}{4\pi G} \frac{D_S}{(D_S - D_L) D_L}
\end{equation}

@@

\section{Cored Isothermal Sphere Profile}
The cored isothermal sphere (CIS) profile is related to the singular isothermal sphere profile, which is often used to describe halos and other collapsed astrophysical objects because of its simple formulation. Unlike the singular isothermal sphere, the cored isothermal sphere does not have a density singularity at its center due to a finite core radius $r_c$. The CIS density profile is defined:

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{isothermalproperties.pdf}
    \caption{Lensing quantities for a CIS profile with $M_{200} = 10^{15} M_\odot$ and $r_c = 10$ kpc.}
    \label{}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{nfwproperties.pdf}
    \caption{Lensing quantities for a NFW profile with $M_{200} = 10^{15} M_\odot$ and $c = 10$.}
    \label{}
\end{figure}

@@

\begin{enumerate}
    \item Consider single foreground lens halo with many background galaxies.
    \begin{itemize}
        \item Start with one NFW halo, then maybe consider more tests.
    \end{itemize}
    \item Construct background galaxies:
    \begin{itemize}
        \item Number density on sky: 50 galaxies/square arcminute (LSST, https://arxiv.org/pdf/1305.0793.pdf)
        \item Intrinsic ellipticity drawn from Gaussian distribution with $\mu=0, \sigma=0.2$ (consistent with https://arxiv.org/pdf/1509.05058.pdf)
        \item Assume they are all that the same distance $D_S$ since this can be determined by redshift (neglecting some noise)
    \end{itemize}
    \item Recommended values:
    \begin{itemize}
        \item $z_L = 0.3$ (bullet cluster)
        \item $z_S = 1.0$ (Hubble Deep Field) (also consistent with https://arxiv.org/pdf/1509.05058.pdf)
        we use the planck 2015 results
        \item $M_{halo} = 10^{15} M_\odot$
    \end{itemize}
\end{enumerate}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]{isothermalellipticities.pdf}
    \includegraphics[width=0.49\textwidth]{nfwellipticities.pdf}
    \caption{plots}
    \label{}
\end{figure}

@@

\section{Results}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{comparison.pdf}
    \caption{plots}
    \label{}
\end{figure}

\section{Conclusions}

@@

has been applied to observations before
https://arxiv.org/pdf/astro-ph/9602053.pdf
but not in the context of comparing density profiles using for strong and weak
but this is highly dependent on good data

http://iopscience.iop.org/article/10.1086/590049/pdf looked at NFW vs isothermal in strong lensing with a focus on arcs, and found differing levels of distinguishibility for different lenses.

https://arxiv.org/pdf/1101.0650.pdf - our results don't match
